---
title: "intro_to_shiny"
author: "Valeriia Vasylieva"
date: '2025-03-03'
---

## Before the class

Ми розробимо додаток, що хоститься локально, і відображає декілька даних із 
статті працівників Антарктичної станції Вернадського. 
Можете почитати статтю, але це не обов'язково.
https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13015 

Для тих, хто захоче зануритися в тему глибже:
* Настільна книга https://mastering-shiny.org/index.html
* Як ділитися своїми застосунками https://shiny.posit.co/r/getstarted/shiny-basics/lesson7/

### Set full working directory

```{r results='hide'}
# For example, "/home/user/Documents/M1S15/"
setwd("./")
```

### Check needed packages

```{r results='hide'}
require(shiny)
require(DT)
requir(rsconnect)
require(ggplot2)
```

### Download and unarchive data in working directory

* Past link in browser
  https://www.ncei.noaa.gov/archive/archive-management-system/OAS/bin/prd/jquery/download/171619.1.1.tar.gz 
* Transfer archive to your working directory
* Unarchive the file





## During the class

### Fill pre-required data

```{r}
data_path0<-"0171619.1.1/0171619/1.1/data/0-data/"
headerphoto<-"NCEI 0171619.jpg"
data_path<-"0171619.1.1/0171619/1.1/data/1-data/"
datasets<-c("","attendance v1-0-DATA.csv","interval v1-0-DATA.csv","repro v1-1-DATA.csv","validation v1-1-DATA.csv")
metadatas<-c("","attendance v1-0-READ ME.txt","interval v1-0- READ ME.txt","repro v1-1-READ ME.txt","validation v1-1-READ ME.txt")
names(metadatas) <- datasets
page.title<-"Estimating nest-level phenology and reproductive success of colonial seabirds using time-lapse cameras"
paper.link<-"https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13015"
```

### Cheet 1 | data

```{r}
# Ensure input is not NULL
  req(input$dataset)  
  # Load data
  df <- read.csv(file.path(wd(),data_path, input$dataset))
  # Drop columns with no values
  df <- df[,colSums(is.na(df))<nrow(df)]
  #Fill NA
  df[is.na(df)] <- "NA"
  # Convert only character columns to factors
  char_cols <- sapply(df, is.character)
  df[char_cols] <- lapply(df[char_cols], as.factor)
  # Convert string to date type
  if ("DATE" %in% names(df)){
    df$DATE<-as.Date(as.character(df$DATE), format = "%d/%m/%Y")
  }
  # Give back the data
  return(df)
```

### Cheet 2 | metadata

```{r}
# Ensure input is not NULL
req(input$dataset)  
# Assemble file path
fileName <- file.path(wd(),data_path, metadatas[input$dataset])  
# Read entire file as text
if (file.exists(fileName)) {
  return(paste(readLines(fileName), collapse = "\n")) 
} else {
  return("Metadata file not found.")
}
```

### Cheet 3 | plot data

```{r}
req(input$upload)
# Get file extention
ext <- tools::file_ext(input$upload$name)
# Validate .csv file extention
switch(ext,
       csv = vroom::vroom(input$upload$datapath, delim = ","),
       validate("Invalid file; Please upload a .csv file")
)
# Load the data
df <- read.csv(paste0(wd(),input$upload$name))
# Transform string to numerical 
df$Latitude <- as.numeric(gsub("−", "-", df$Latitude))
df$Longitude <- as.numeric(gsub("−", "-", df$Longitude))
# Give back the data
return(df)
```

### Cheet 4 | plot

```{r}
ggplot(species_locality_data(), aes(x = Longitude, y = Latitude, shape = Species, color = Species)) +
  geom_point(size = 4, position = "jitter") +  # Adjust point size
  scale_shape_manual(values = c("Adélie" = 16, "Gentoo" = 17, "Chinstrap" = 15)) +
  scale_color_manual(values = c("Adélie" = "black", "Gentoo" = "red", "Chinstrap" = "blue")) +
  theme_minimal() +
  labs(title = "Penguin Localities", x = "Longitude", y = "Latitude") +
  theme(legend.title = element_blank())
```


